Generalized Tractability

### 2023-04-25 GeneralizedTractability_230418.tex
Updates from Overleaf

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation} \label{eq:eigenvalues}
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation}
\fred{Are we restricting ourselves by not allowing $\lambda_{1,d} \to \infty$ as $d \to \infty$}
This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.  The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering.

@@

@@

\fred{
\begin{itemize}
\item Do we consider the case of normalized error? (Peter) 
\item Are there cases of $T$ that do not correspond to known tractability measures? (Kachi)
\item Restricted domain? (Fred)
\end{itemize}}

>>>>

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation} \label{eq:eigenvalues}
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation}
\fred{Are we restricting ourselves by not allowing $\lambda_{1,d} \to \infty$ as $d \to \infty$?}
\peter{Yes, indeed, I think we do (e.g., think of tensor product spaces where the first eigenvalue is bigger than one). But I am wondering whether we could do without $\lambda_{\max}$ by the assumption that $T$ is bounded away from zero by some absolute value $t$. Then, in all cases we could bound $1/T(\cdot,\cdot,\cdot)$ by $1/t$. Actually, is $T(0,1,\boldsymbol{0})$ our $t$?}

This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.  The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering.

@@

\section{Try Restricted Region}

Suppose that the domain of interesting $(\varepsilon^{-1},d)$ is given by 
\begin{equation}
    \Omega  = \{(\varepsilon^{-1},d) : \varepsilon > E(d) \ge 0, \ d \in \naturals \},
\end{equation}
where $E$ is a non-increasing function.  Furthermore, for all $d \in \naturals$, let $N(d) = \min \{n : \lambda_{n+1,d} \le E(d) \}$.  This means that 
\[
\lambda_{1,d}^{-1} \le \cdots \le \lambda_{N(d),d}^{-1} < E(d) \le \lambda_{N(d)+1,d}^{-1}, \qquad (\lambda_{1,d}^{-1},d), \ldots, (\lambda_{N(d),d}^{-1},d) \in \Omega.
\]

\begin{theorem}\label{thm_main_strong_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^{N(d)} \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{strct}} = \widetilde{\cp}_{\textup{strct}}$, and the set of optimal strong tractability parameters is
\[
	\cp_{\textup{sopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{strct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{strct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d',\vp)$ are non-decreasing in $i$ for any $d'\in\{1, \ldots, d\}$. In particular, for all $n\in \{1, \ldots, N(d')\}$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d',\vp)} \le \cdots \le \frac{1}{T(\lambda_{1, d}^{-1},d',\vp)} \\
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp) }
    \le \frac 1n \sum_{i=1}^n  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) }
    \le \frac 1n \sum_{i=1}^{N(d')}  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n \ge T(\varepsilon^{-1},d',\vp) \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le
   \frac 1n \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) } \le \frac{1}{T(\varepsilon^{-1},d',\vp)} \qquad \forall \varepsilon > 0, \ d,d' \in \naturals.
\end{multline}

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
    \begin{align} %\label{eq:comp_equiv}
    \nonumber
    \comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
    \nonumber
    & = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},d',\vp)}\le \frac{1}{T(\varepsilon^{-1},d',\vp) }\biggr\} \qquad \forall d' \in \naturals\\
    \nonumber
    & \le  \min \biggl\{n \in \natzero : n \ge T(\varepsilon^{-1},d',\vp) \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \biggr\} \quad \forall d' \in \naturals \qquad \text{by \eqref{eq:implicate_a}} \\
    & \le  T(\varepsilon^{-1},d',\vp) \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} + 1 \quad \forall d' \in \naturals. \label{eq:compUP}
\end{align}

@@

\fred{
\begin{itemize}
\item Do we consider the case of normalized error? \peter{What would most likely change then, is: we would normalize all eigenvalues by dividing by the values of the $\lambda_{1,d}$, and all summations in the conditions could start with the index 1. It might be a bit more tricky for sub-$h$ tractability, where we would maybe make sure that all functions are still suitably defined. However, I would not expect new proof ideas. Maybe a remark along the lines of ``presumably everything works analogously'' is sufficient?}
\item Are there cases of $T$ that do not correspond to known tractability measures? (Kachi)
\item Restricted domain? (Fred)
\end{itemize}}
